### 2016-12-12 Projects/DelCo_2012/Report/DelCo_2012.tex
Minor edits to dimension accuracy and acknowledgement section

The fire suppression series included 20 experiments. Twelve experiments were conducted in the single-story structures (six in each structure) using both water and CAF application as a suppression agent. The nozzle patterns were varied and included straight stream, solid stream, and narrow fog stream. Six experiments were conducted in the lower level of the two-story structure using both water and CAF from a straight stream nozzle. Additionally, two CAF experiments were conducted in the two-story burn building with a solid stream. The lower level of the two-story structure allowed for a fire compartment that was approximately twice as large as the fire compartment in the single-story.

@@

The burn building, or live fire training structure, contained a two-story section and a three-story section. The building was supported with reinforced concrete beams and columns. The floors and ceilings were also concrete, and the interior walls were constructed of cement block. The walls and ceiling of the burn rooms were protected with a 25~mm (1~in) thick layer of calcium silicate insulation and was covered by 50~mm (2~in) thick concrete tile. The floors of the burn rooms were protected with fire brick. Tests were conducted in a two-room configuration located on the ground floor of the two-story section. The overall dimensions of the test floor were 6.5~m (21.3~ft) by 9.2~m (30.2~ft). Figure~\ref{fig:Delaware_County,_PA_Burn_Building_Layout} contains a basic plan view of the two-room configuration, and a fully dimensioned plan view of the test floor is presented in Figure~\ref{fig:Delaware_County,_PA_Burn_Building_Layout_Detailed} in Appendix~\ref{app:floor_plans}.

@@

The fuel load was located in the burn room, which had interior dimensions of 3.8~m (12.5~ft) by 5.7~m (18.7~ft) and a ceiling height of 3.4~m (11.0~ft). Water application for the purpose of gas cooling always occurred in the adjacent room, which had interior dimensions of 4.1~m (13.4~ft) by 5.7~m (18.7~ft) and a ceiling height of 3.4~m (11.0~ft). The open doorway that connected the burn room to the gas cooling room and the doorway from the gas cooling room to the exterior both measured 2.0~m (6.5~ft) high and 0.9~m (3.0~ft) wide.

@@

Two identical single-story concrete structures were built on a concrete slab as shown in Figure~\ref{fig:Delaware_County,_PA_Fire_Test_Structures}. They were designed to simulate a single floor of a residential structure. The outer wall of each structure was composed of interlocking concrete blocks 0.6~m (2.0~ft) wide, 0.6~m (2.0~ft) high and 1.2~m (4.0~ft) long. The interior dimensions of each structure were 6.1~m (20.0~ft) wide, 11.0~m (36.0~ft) long, and 2.4~m (8.0~ft) high. The joints and gaps between the blocks were filled with high temperature insulation.

@@

Each structure had a burn room with interior dimensions of approximately 6.0~m (19.6~ft) by 4.4~m (14.3~ft) by 2.4~m (8.0~ft) high. A 5.4~m (17.6~ft) long hallway with a width of 0.9~m (3.1~ft) and ceiling height of 2.4~m (8.0~ft) connected the burn room to a small entry foyer at the front of the structure. The entry foyer was 1.2~m (4.1~ft) by 1.9~m (6.1~ft) and had a ceiling height of 2.4~m (8.0~ft). A 2.0~m (6.5~ft) high by 0.9~m (3.0~ft) wide open doorway on the north side of the structure connected the entry foyer to the exterior. The opening on the south face of the structure from the exterior to the burn room was 2.4~m (7.75~ft) high and 1.09~m (3.6~ft) wide. A basic plan view of the structure is presented in Figure~\ref{fig:Test_Structure_Floor_Plan}, and a fully dimensioned plan view is included in Figure~\ref{fig:Test_Structure_Floor_Plan_Detailed} in Appendix~\ref{app:floor_plans}.

@@

The floor of the structure was the concrete pad. The interior walls of the structure were framed with steel studs and track. The studs were set to 0.4~m (16~in) centers. The ceiling support was composed of wood truss joist I-beams (TJIs) with a 0.3~m (11.75~in) depth. Each TJI was composed of laminated veneer lumber flanges with a cross section of 29~mm (1.13~in) x 44~mm (1.75~in) and an 11~mm (0.43~in) thick oriented strand board (OSB) web. Tongue and grove, 18~mm (0.72~in) thick, OSB was attached to the top of the TJIs.

@@

A set of stairs, each with a 180~mm (7.25~in) rise and 190~mm (7.5~in) run, was built along the interior of the east wall to connect the two floors of the structure. The stairway was 1.2~m (4.0~ft) wide, and the edge of the bottom step was located 1.6~m (5.25~ft) from the south wall.The second story walls were wood framed with 51~mm (2.0~in) by 102~mm (4.0~in) studs. The studs were set to 0.40~m (16~in) centers. The interior walls were protected by a layer of 16~mm (0.63~in) fire rated gypsum board, 16~mm (0.63~in) Durock cement board, and a second layer of 16~mm (0.63~in) fire rated gypsum board. The exterior walls were protected by 11~mm (0.31~in) thick OSB and 8~mm (0.31~in) thick fiber cement lap siding. Figure~\ref{fig:dimensioned_second_2story} shows a basic plan view of the second story of the modified west structure, and a fully dimensioned plan view of the floor is provided in Figure~\ref{fig:dimensioned_second_2story_detailed} in Appendix~\ref{app:floor_plans}.

@@

The structures were instrumented for gas temperature, gas velocity, and heat flux measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single, sheathed thermocouples were installed in conjunction with the bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples with a 1.0~mm (0.04~in) nominal diameter. The thermocouple was protected with an 3.2~mm (0.13~in) diameter inconel sheath. The sheathing protects the thermocouple from water but slows the response time because of the additional thermal mass. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium plate to prevent contributions from convective heat transfer. A legend to clarify the instrumentation schematics presented in the following sections is included in Figure~\ref{fig:Instrumentation_Legend}.

@@

There were seven suppression tests conducted in the two-story, purpose-built structure. All seven of the tests used a fuel load consisting of three sofas, polypropylene carpeting over polyurethane foam padding, and OSB paneling along three walls and the ceiling. Five of the seven tests also had two stacks of wooden pallets in addition to the initial fuel load (Section~\ref{sec:fire_suppression_furniture_fuel}). Ignition occurred in the southwest corner of the ground floor, and available ventilation was through the ground floor double doors and the south exterior door on the second story of the structure. The fires were intended to replicate room and contents basement fires typically seen within residential structures. Suppression occurred at the ground level double doors of the structure. Table~\ref{tab:Test_Descriptions_2} provides an overview of the suppression tests conducted.

@@

Seven of the suppression experiments were conducted in a two-story structure.  The fire compartment was the full lower level of the structure, and it had an open floor plan with an area of 51.4~m$^2$ (553~ft$^2$). At one end of the room was an open doorway that was 1.8~m (6.0~ft) wide and 2.0~m (6.7~ft) high and led to a stairwell. A manual hose stream application was made from outside the doorway on the lower level. In these experiments, there was a combination of wood-based and synthetic fuels. The fires were allowed to transition through flashover before the initial fire suppression attack was made. In each case, the initial knockdown of the fire required no more than 60~gal of plain water or water in the form of CAF.

@@

This study was part of a larger DHS/FEMA Assistance to Firefighters Research and Development Grant project led by Cal Poly. The authors would like to acknowledge the support of the Cal Poly team, especially: Fred Mowrer, Christopher Dicus, Thomas Korman, and Sean Mitchell.

@@

Thanks are given to the dedicated team of fire instructors and firefighters that supported these experiments; Brian Righter (Rudy), John McGowan, Matthew Poissant, John Frey, Brian Embert, Raymond Keller III, Mike Bramble, Edward McBride, Mark Morrissey, Bill Norris, and Matt LeTourneau.

>>>>

The fire suppression series included 20 experiments. Twelve experiments were conducted in the single-story structures (six in each structure) using both water and CAF application as a suppression agent. The nozzle patterns were varied and included straight stream, solid stream, and narrow fog stream. Six experiments were conducted in the lower level of the two-story structure using both water and CAF from a straight stream nozzle. Additionally, two CAF experiments were conducted in the two-story structure with a solid stream. The lower level of the two-story structure allowed for a fire compartment that was approximately twice as large as the fire compartment in the single-story.

@@

The burn building, or live fire training structure, contained a two-story section and a three-story section. The building was supported with reinforced concrete beams and columns. The floors and ceilings were also concrete, and the interior walls were constructed of cement block. The walls and ceiling of the burn rooms were protected with a 25~mm (1.0~in) thick layer of calcium silicate insulation and was covered by 51~mm (2.0~in) thick concrete tile. The floors of the burn rooms were protected with fire brick. Tests were conducted in a two-room configuration located on the ground floor of the two-story section. The overall dimensions of the test floor were 6.5~m (21.3~ft) by 9.2~m (30.2~ft). Figure~\ref{fig:Delaware_County,_PA_Burn_Building_Layout} contains a basic plan view of the two-room configuration, and a fully dimensioned plan view of the test floor is presented in Figure~\ref{fig:Delaware_County,_PA_Burn_Building_Layout_Detailed} in Appendix~\ref{app:floor_plans}.

@@

The fuel load was located in the burn room, which had interior dimensions of 3.8~m (12.5~ft) by 5.7~m (18.7~ft) and a ceiling height of 3.4~m (11.0~ft). Water application for the purpose of gas cooling always occurred in the adjacent room, which had interior dimensions of 4.1~m (13.4~ft) by 5.7~m (18.7~ft) and a ceiling height of 3.4~m (11.0~ft). The window on the west face of the structure from the exterior to the burn room was 1.1~m (3.5~ft) by 1.1~m (3.5~ft).

@@

Two identical single-story concrete structures were built on a concrete slab as shown in Figure~\ref{fig:Delaware_County,_PA_Fire_Test_Structures}. They were designed to simulate a single floor of a residential structure. The outer wall of each structure was composed of interlocking concrete blocks 0.6~m (2.0~ft) wide, 0.6~m (2.0~ft) high, and 1.2~m (4.0~ft) long. The interior dimensions of each structure were 6.1~m (20.0~ft) wide, 11.0~m (36.0~ft) long, and 2.4~m (8.0~ft) high. The joints and gaps between the blocks were filled with high temperature insulation.

@@

Each structure had a burn room with interior dimensions of approximately 6.0~m (19.6~ft) by 4.4~m (14.3~ft) by 2.4~m (8.0~ft) high. A 5.4~m (17.6~ft) long hallway with a width of 0.9~m (3.1~ft) and ceiling height of 2.4~m (8.0~ft) connected the burn room to a small entry foyer at the front of the structure. The entry foyer was 1.2~m (4.1~ft) by 1.9~m (6.1~ft) and had a ceiling height of 2.4~m (8.0~ft). A 2.0~m (6.5~ft) high by 0.9~m (3.0~ft) wide open doorway on the north side of the structure connected the entry foyer to the exterior. The window on the south face of the structure from the exterior to the burn room was 1.2~m (4.0~ft) tall by 1.1~m (3.5~ft) wide. A basic plan view of the structure is presented in Figure~\ref{fig:Test_Structure_Floor_Plan}, and a fully dimensioned plan view is included in Figure~\ref{fig:Test_Structure_Floor_Plan_Detailed} in Appendix~\ref{app:floor_plans}.

@@

The floor of the structure was the concrete pad. The interior walls of the structure were framed with steel studs and track. The studs were set to 0.4~m (16.0~in) centers. The ceiling support was composed of wood truss joist I-beams (TJIs) with a 0.3~m (11.8~in) depth. Each TJI was composed of laminated veneer lumber flanges with a cross section of 29~mm (1.1~in) x 44~mm (1.8~in) and an 11~mm (0.43~in) thick oriented strand board (OSB) web. Tongue and grove, 18~mm (0.72~in) thick, OSB was attached to the top of the TJIs.

@@

A set of stairs, each with a 180~mm (7.25~in) rise and 190~mm (7.5~in) run, was built along the interior of the east wall to connect the two floors of the structure. The stairway was 1.2~m (4.0~ft) wide, and the edge of the bottom step was located 1.6~m (5.25~ft) from the south wall.The second story walls were wood framed with 51~mm (2.0~in) by 102~mm (4.0~in) studs. The studs were set to 0.4~m (16.0~in) centers. The interior walls were protected by a layer of 16~mm (0.63~in) fire rated gypsum board, 16~mm (0.63~in) Durock cement board, and a second layer of 16~mm (0.63~in) fire rated gypsum board. The exterior walls were protected by 11~mm (0.31~in) thick OSB and 8~mm (0.31~in) thick fiber cement lap siding. Figure~\ref{fig:dimensioned_second_2story} shows a basic plan view of the second story of the modified west structure, and a fully dimensioned plan view of the floor is provided in Figure~\ref{fig:dimensioned_second_2story_detailed} in Appendix~\ref{app:floor_plans}.

@@

The structures were instrumented for gas temperature, gas velocity, and heat flux measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single, sheathed thermocouples were installed in conjunction with the bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples with a 1.0~mm (0.04~in) nominal diameter. The thermocouple was protected with a 3.2~mm (0.13~in) diameter inconel sheath. The sheathing protects the thermocouple from water but slows the response time because of the additional thermal mass. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium plate to prevent contributions from convective heat transfer. A legend to clarify the instrumentation schematics presented in the following sections is included in Figure~\ref{fig:Instrumentation_Legend}.

@@

There were eight suppression tests conducted in the two-story, purpose-built structure. All eight of the tests used a fuel load consisting of three sofas, polypropylene carpeting over polyurethane foam padding, and OSB paneling along three walls and the ceiling. Six of the eight tests also had two stacks of wooden pallets in addition to the initial fuel load (Section~\ref{sec:fire_suppression_furniture_fuel}). Ignition occurred in the southwest corner of the ground floor, and available ventilation was through the ground floor double doors and the south exterior door on the second story of the structure. The fires were intended to replicate room and contents basement fires typically seen within residential structures. Suppression occurred at the ground level double doors of the structure. Table~\ref{tab:Test_Descriptions_2} provides an overview of the suppression tests conducted.

@@

Eight of the suppression experiments were conducted in a two-story structure.  The fire compartment was the full lower level of the structure, and it had an open floor plan with an area of 51.4~m$^2$ (553~ft$^2$). At one end of the room was an open doorway that was 1.8~m (6.0~ft) wide and 2.0~m (6.7~ft) high and led to a stairwell. A manual hose stream application was made from outside the doorway on the lower level. In these experiments, there was a combination of wood-based and synthetic fuels. The fires were allowed to transition through flashover before the initial fire suppression attack was made. In each case, the initial knockdown of the fire required no more than 60~gal of plain water or water in the form of CAF.

@@

This study was part of a larger DHS/FEMA Assistance to Firefighters Research and Development Grant project led by Cal Poly. The authors would like to acknowledge the support of the Cal Poly team, especially Fred Mowrer, Christopher Dicus, Thomas Korman, and Sean Mitchell.

@@

Thanks are given to the dedicated team of fire instructors and firefighters that supported these experiments: Brian Righter (Rudy), John McGowan, Matthew Poissant, John Frey, Brian Embert, Raymond Keller III, Mike Bramble, Edward McBride, Mark Morrissey, Bill Norris, and Matt LeTourneau.
